---
title: "Giant Seabass Movement at Santa Barbara Island"
author: "Ryan Freedman"
date: "11/29/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Add in packages and clean data in this chunk. Data is outputs from VEMCO's VUE software plus some spreadsheets maintatined by CINMS GIS Czar. This includes detections from VUE, spreadsheet of reciever sites around SBI and a list of fish tagged for this project.
```{r}
#Add Packages
library(tidyverse)
library(lubridate)
library(sf)
library(sp)
library(ggpubr)
library(readxl)
library(lubridate)
library(raster)
library(rworldmap)
library(ggplot2)
library(scales) 
library(here)
library(ggrepel)
library(tidyselect)
library(dplyr)
library(ADePTR)
library(raster)
library(gganimate)
library(VTrack)
library(forecast)
library(ggthemes)
library(ATT)
library(VTrack)
library(mapview)
library(rgdal)
library(raster)
library(viridis) ## ac
library(reshape) # this is for colsplit
library(fields) # this is for colorramps
library(AssFunc) # for the abacus plots
library(leaflet)
library(dplyr)
library(ggsn)
library(ggmap)
library(gdistance)
library(asnipe) #for GMM events
library(sna) #for calculating weighted degrees
library(igraph) #provides a set of data types and functions for implementation of graph algorithms and
#fast handling of large graphs, with millions of vertices and edges
library(extrafont)
```

# Add Needed Data

```{r}

#Cleaned Up Detection File (Station Names made to match)
detect <- read_csv("VUE_Export.csv")
detect <- detect %>% dplyr::select(`Date and Time (UTC)`,Transmitter, Receiver,           `Station Name`)  #drop not needed cols
head(detect)  

#VR2W Site Information
sites <- read_csv("GSBsites.csv") #bring in VR2W data
sites<-sites %>% dplyr::select (c("Site","Lat_dd","Lon_dd")) #drop not needed cols
names(sites)[names(sites) == "Site"] <- "Station Name" #rename column for merge
head(sites)

#GSB Data
Tagged_GSB<- read_excel("Tagged GSB (In-House).xlsx") #bring in tag data from notebooks
Tagged_GSB$ID<- paste("A69-1602-", Tagged_GSB$ID, sep="") #ad in tag leader info
my.tags<-Tagged_GSB$ID #make a list of my tag IDs

```

Adding in Spatial Data here. Files include a shapefile of Santa Barbara Island and a delinating project area shapefile. Both are required by the ADePTR package which will come later. Both are exports from ArcGIS Pro.

```{r}
SBI <- st_read(dsn = "SBI_Shapefiles", layer = "SBI")

SBI2<-st_cast(SBI,"POLYGON")
SBI3<-SBI2
SBI4<-SBI3$geometry[-66]
SBI5<-st_cast(SBI4,"POLYGON")

ProjectArea<-st_read(dsn = "SBI_Shapefiles", layer = "SBI_square")
```

Add in Tagging Data from the Notebook

```{r}
# Merge dataset and just get my tags

SBI.dat<-merge(detect, sites, by="Station Name") #Merge detection data with site data
SBI.GSB <-SBI.dat[SBI.dat$Transmitter %in% my.tags, ] # Subset with NOAA Tagged Fish
SBI.GSB$`Date and Time (UTC)`<-as.POSIXct(SBI.GSB$`Date and Time (UTC)`, tz = "UTC",
                                          format="%m/%d/%y %H:%M")

```


Breakdown of Detections by Fish
```{r}
#Get Number of Detections by Transmitter

trans_counts <-SBI.GSB %>%
  dplyr::select (-c(`Date and Time (UTC)`)) %>%
  group_by(Transmitter) %>%
  count(Transmitter) %>%
  ungroup()

trans_counts
```


Time Abacus by Location
```{r}

all <-ggplot(SBI.GSB)+ geom_point(aes(x=`Date and Time (UTC)`, y= `Station Name`, color= Transmitter), shape=15, size=3) + scale_color_manual(values=c("#000033","#003A2F","#01635B","#35918C","#7DC7BD","#D8BB77","#B97D2D","#874E08","#502F06", "#00CC99", "#003000", "#99CCCC"))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_datetime() +
  ggtitle("All Fish") +
  xlab("Month") +
  ylab("Station") +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=15, face= "bold"))

all


```

Time Abacus by Fish
```{r}

ball <-ggplot(SBI.GSB)+ geom_point(aes(x=`Date and Time (UTC)`, y= `Transmitter`, color= Transmitter), shape=15, size=3) + scale_colour_brewer(palette="Spectral")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_datetime() +
  ggtitle("All Fish") +
  xlab("Month") +
  ylab("Station") +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=15, face= "bold"))

ball

```


Data Processing Work With ADePTR ###THIS IS WHERE PIKE IS STUCK
```{r}
#### Data Drop in

rec.log<-read.csv("GSBRecLogtillWinter19.csv")

#make Station File
stations<-SBI.dat %>%
  dplyr::select(sta_id = "Station Name", rec_id = "Receiver",x="Lat_dd", y= "Lon_dd")  %>%
  unique()
xx<-merge(rec.log, stations, by= 'sta_id')
AD.stations<-data.frame(xx$sta_id, xx$rec_id.y, xx$dt_dep, xx$dt_ret, xx$x, xx$y)
names(AD.stations)<-c("sta_id", "rec_id", "dt_dep", "dt_ret", "y", "x")
AD.stations$dt_dep<-as.POSIXct(AD.stations$dt_dep, tz="America/Los_Angeles", format= "%m/%d/%y %H:%M")
AD.stations$dt_ret<-as.POSIXct(AD.stations$dt_ret, tz="America/Los_Angeles", format= "%m/%d/%y %H:%M")

#Make Detection File
AD.detections<-SBI.dat %>%
  dplyr::select(id = "Transmitter", dt = "Date and Time (UTC)", rec_id="Receiver")

AD.detections$dt<-as.POSIXct(AD.detections$dt, tz="America/Los_Angeles", format="%m/%d/%y %H:%M")

#Process detections and stations
proc.det <- proc_dets(det = AD.detections, sta = AD.stations, crs = 4326)

```

Working through Analysis with ADEPTR

Speed Filter
```{r}
filt.det <- proc.det %>% 
  spd_filter(min_dist = units::set_units(1, "mi"))  %>%  #Speed filter 
  singleton_filter(time_horizon = "1 week") #singleton filter
```

Save "fit.det" as a CSV in the SBI_GSB working directory
```{r}
write.csv(filt.det, "filter_dections_all.csv")
```

```{r}
plot_sta_history(proc_det = filt.det) #dections by station

p <- map_dets(proc_det = filt.det, use_ggplot = TRUE)$ggplot
p+geom_sf(data = SBI5, fill="darkgreen")


```

Periodicity 
```{r}
tsData <- filt.det$dt # ts data

tms <- as.POSIXct(tsData, format="%H:%M:%S")
hours <- substr(tms, 12, 16)
hours<-as.POSIXct(hours, format="%H:%M")
hist( hours , breaks = unique( hours ) )

# Rerun for Mound Offshore
ts.MO<-filt.det %>% 
  subset(sta_id =="Mound Offshore")
hours.MO <- substr(ts.MO$dt, 12, 16)
hours.MO<-as.POSIXct(hours.MO, format="%H:%M")
hist( hours.MO , breaks = unique( hours.MO ) )

```

Daily Averages for Mound Offshore Site for comparison with ST
```{r}
day.filt.det<-filt.det
day.filt.det$Date <- substr(day.filt.det$dt, 1, 10)

day.filt.det2<- day.filt.det %>% group_by(Date) %>% summarise(frequency = n())
day.filt.det2<-st_drop_geometry(day.filt.det2)
day.filt.det2$Year<-substr(day.filt.det2$Date, 1, 4)
day.filt.det2$Date2<-as.POSIXct(day.filt.det2$Date, format="%Y-%m-%d")

ggplot(day.filt.det2, aes(x = Date2, y= frequency,fill=Year)) + geom_bar(stat="identity") + ylab("Date") + xlab("Detections") + theme_tufte()

ggplot(day.filt.det2, aes(x = Date2, y= frequency,fill=Year)) + geom_bar(stat="identity") +
  scale_x_datetime(limits = as.POSIXct(c("2018-08-13", "2018-12-31"))) +ylab("Date") +   xlab("Detections")+ theme_tufte()
  
 


#SoundTrap Site
ts.MO$Date <- substr(ts.MO$dt, 1, 10)
ST.data<- ts.MO %>% group_by(Date) %>% summarise(frequency = n())
ST.data<-st_drop_geometry(ST.data)

ST.data$Year<-substr(ST.data$Date, 1, 4)
ST.data$MonDay<-substr(ST.data$Date, 6, 10)
ST.data$Date2<-as.POSIXct(ST.data$Date, format="%Y-%m-%d")

#Plot.
ggplot(ST.data, aes(x = Date2, y=frequency, fill = Year)) + geom_bar(stat="identity") +ylab("Date") +   xlab("Detections") + theme_tufte()
  

```


Time Series
```{r}
decomposedRes <- decompose(tsData) # use type = "additive" for additive components
plot (decomposedRes) # see plot below
stlRes <- stl(tsData, s.window = "periodic")

```


#VTRACK Packages

VTRACK Package Starts Here. Lets begin by reorging data to the right format. Vtrack is very sensitive to data formatting and requires data in a specific format. VTrack will throw errors without data in the correct format

##Data QA/QC
```{r}
#Distances File

SBI.distance<-sites %>% transmute(
  LOCATION = `Station Name`,
  LATITUDE = Lat_dd,
  LONGITUDE = Lon_dd,
  RADIUS = 200)

SBI.distance$LOCATION<-as.factor(SBI.distance$LOCATION)
SBI.distance$RADIUS<-as.integer(SBI.distance$RADIUS)
SBI.distance<-as.data.frame(SBI.distance)

DirectDM <- GenerateDirectDistance(SBI.distance)

#Tagging Info

sbi.taginfo<-Tagged_GSB %>% transmute(
  transmitter_id = ID,
  tag_id = substring(Tagged_GSB$ID, 10),
  release_id = "",
  tag_project_name = "GSB Soundscapes",
  scientific_name = "Stereolipas gigas",
  common_name = "Giant Seabass",
  embargo_date = NA,
  is_protected = "Yes",
  release_latitude = LATITUDE,
  release_longitude = LONGITUDE,
  ReleaseDate = DATE,
  sensor_slope = NA,
  sensor_intercept = NA,
  sensor_type = "PINGER",
  sensor_unit = NA,
  tag_model_name = "",
  tag_serial_number = "",
  tag_expected_life_time_days =653,
  tag_status = "",
  sex = "",
  measurement = "",
  dual_sensor_tag = 'FALSE'
)

sbi.taginfo <- filter(sbi.taginfo, tag_id != "16716")

#station info

SBI.statinfo<-AD.stations %>% transmute(
  project_name = "GSB Soundscapes",
  installation_name = "Santa Barbara Island",
  station_name = sta_id,
  receiver_name = rec_id,
  deploymentdatetime_timestamp = dt_dep,
  recoverydatetime_timestamp = dt_ret,
  status = "RECOVERED",
  station_longitude = x,
  station_latitude = y,
  imos_device = "FALSE",
  region = "California Counter Current")


# Detection Data

v.data<-SBI.GSB %>% transmute(
  Date.and.Time..UTC. = `Date and Time (UTC)`,
  Receiver = Receiver,
  Transmitter.Name = substring(Transmitter, 10), 
  Transmitter = Transmitter,
  Transmitter.Serial = "",
  Sensor.Value = NA,
  Sensor.Unit = NA,
  Station.Name = `Station Name`,
  Latitude = Lat_dd,
  Longitude = Lon_dd,
  Ascention = "",
  Year = "",
  region = "California Counter Current",
  posix = as.POSIXct(`Date and Time (UTC)`),
  local = as.POSIXct(`Date and Time (UTC)`),
  station.short = `Station Name`,
  Date = as.factor(substring(`Date and Time (UTC)`, 1,10))
)

attributes(v.data$local)$tzone <- "America/Los_Angeles"

## this is necessary for setting up the data and calculating the centers of activity ###

v.data$Latitude<-as.numeric(v.data$Latitude)
v.data$Longitude<-as.numeric(v.data$Longitude)
v.data$Transmitter<-as.factor(v.data$Transmitter)
v.data$Station.Name<-as.factor(v.data$Station.Name)
v.data$Receiver<-as.factor(v.data$Receiver)

#Drop Fish 161716 Not Enough Detections
v.data <- filter(v.data, Transmitter.Name != "16716")
```

This is the Animal Tracking Tool Box Function and the way to store data in a single object
```{r}

## Setup data for use with the Animal Tracking Toolbox
ATTdata.SBI<- setupData(Tag.Detections = v.data,
                      Tag.Metadata = sbi.taginfo, 
                      Station.Information = SBI.statinfo, 
                      source="VEMCO")


ATTdets.SBI<-as.data.frame(ATTdata.SBI$Tag.Detections)
ATTmeta.SBI<-as.data.frame(ATTdata.SBI$Tag.Metadata)
```

### Make abacus plots of raw detections ###
```{r}
abacusPlot(ATTdata.SBI)
abacusPlot(ATTdata.SBI, facet=T, new.window = T)
```

```{r}
## Calculate detection metrics
detSum.SBI<- detectionSummary(ATTdata.SBI)

detSum.SBI_Overall<-as.data.frame(detSum.SBI$Overall)

detSum.SBI_Subsetted<-as.data.frame(detSum.SBI$Subsetted)
```


##Core Areas of Activity

This chunk develops hourly Core Areas of Activity for each fish. It's very process heavy and takes awhile to think. It took a whole 12 hour day for 2 years of data

```{r}
## Estimate Short-term Center of Activities for all individuals

COAdata<-ATT::COA(ATTdata.SBI, 
             timestep = 60, ## timestep bin used to estimate centers of activity (in minutes)
              split = TRUE)

COAdata

## Define coordinate system for projection of detection data (in m)
proj <- CRS("+init=epsg:3857")

## Estimate 20%, 50% and 95% BBKUD contour areas and store polygons
BBkud_est.SBI <- HRSummary(COAdata, 
                         projCRS=proj, 
                         type="BBKUD", 
                         cont=c(50,95), 
                         storepoly=TRUE, h = 200,
                         ext = 5, grid = 1000, div = 2)


## Select rasters of full KUDs for each individual into a single list
fullstack.SBI <-
  unlist(BBkud_est.SBI$Spatial.Objects)[grep("*_full", names(unlist(BBkud_est.SBI$Spatial.Objects)))]

names(fullstack.SBI) <-
  unlist(lapply(strsplit(names(fullstack.SBI), "[.]"), `[[`, 1))


```

Plot the Data in the Leaflet package for all KUDs together. Leaflet allows for a google map like exploration of the KUDs

```{r}

library(leaflet)

## Full KUD for all tagged animals
fullmap <- leaflet() %>%
  addProviderTiles(providers$Esri.WorldImagery)

for (i in 1:length(fullstack.SBI)) {
  tempras<-disaggregate(fullstack.SBI[[i]], fact=3, method='bilinear')
  values(tempras)[values(tempras) >95] <-NA
  fullmap <- 
    fullmap %>% 
    addRasterImage(tempras, opacity = 0.8, group = names(fullstack.SBI)[i])
}

coa.detections<-
  do.call(rbind, COAdata) 

t.coa.detections <- as.data.frame(t(coa.detections)) #name change here incase code 

fullmap <- 
  fullmap %>%
  addCircleMarkers(lng = coa.detections$Longitude.coa, lat = coa.detections$Latitude.coa,
                   color = "red", radius = 1, weight=1, group = coa.detections$Tag.ID) %>%
  addCircleMarkers(lng = statinfo$station_longitude, lat = statinfo$station_latitude,
                   fill = F, color = "white", radius = 4, weight = 2, group = "Receiver Stations") %>%
  addMeasure(position = "bottomleft",
             primaryLengthUnit = "meters",
             primaryAreaUnit = "sqmeters") %>%
  addLayersControl(
    baseGroups = coa.detections$Tag.ID,
    overlayGroups = "Receiver Stations",
    options = layersControlOptions(collapsed = FALSE)
  )

fullmap
```

Select rasters of subsetted KUDs for each individual into a single list

```{r}
substack <-
  unlist(BBkud_est.SBI$Spatial.Objects)[grep("*_sub", names(unlist(BBkud_est.SBI$Spatial.Objects)))]
```

This code singles out a particular tag but requires the user to change the tag codepace according to your dataset. Need to debug
```{r}
tag_subset <-
  disaggregate(substack$`9718.BBKUD_sub`,
               fact = 3,
               method = 'bilinear')

values(tag_subset)[values(tag_subset) > 96] <- NA
names(tag_subset) <- 
  format(ymd(paste0(substring(
    names(tag_subset), 2
  ), ".01")), "%B %Y")

tag.det<-
  coa.detections %>%
  mutate(sub = format(TimeStep.coa, "%B.%Y"))

coa.datr<-
  ungroup(tag.det) %>%
  dplyr::select(Longitude.coa,Latitude.coa,sub) %>%
  split(.$sub)


submap<- 
  leaflet() %>%
  addProviderTiles(providers$Esri.WorldImagery)

for (i in 1:length(names(tag_subset))) {
  submap <- 
    submap %>% 
    addRasterImage(tag_subset[[i]], opacity = 0.5, group = gsub("[.]", " ", names(tag_subset)[i]))
} 

submap<- 
  submap %>%
  addCircleMarkers(lng = tag.det$Longitude.coa, lat = tag.det$Latitude.coa,
                   color = "red", radius = 1, weight=1, group = gsub("[.]", " ", tag.det$sub)) %>%
  addCircleMarkers(lng = sites$Lat_dd, lat = sites$Lon_dd,
                   fill = F, color = "white", radius = 4, weight=2, group = "Station Name") %>%
  addLayersControl(
    baseGroups = gsub("[.]", " ", names(tag_subset)),
    overlayGroups = "Receiver Stations",
    options = layersControlOptions(collapsed = FALSE)
  )

submap
```

##VTrack Animation!: Multitag beep plot
```{r}
v.data.like.crocs<-v.data %>% transmute(
  Date.Time = Date.and.Time..UTC.,
  Code.Space = substring(Transmitter, 1, 8),
  ID = substring(Transmitter, 10), 
  Sensor.1 = NA,
  Units.1 = NA,
  Sensor.2 = NA,
  Units.2= NA,
  Transmitter.Name = substring(Transmitter, 10),
  Transmitter.S.N = substring(Transmitter, 10),
  Receiver.Name = Receiver,
  Receiver.S.N. = substring(Receiver, 6),
  Station.Name = Station.Name,
  STATIONNAME = Station.Name,
  Station.Latitude = Latitude,
  Station.Longitude = Longitude,
)

SBI.distance.like.crocs<-SBI.distance %>% transmute(
  STATIONNAME = LOCATION,
  LATITUDE = LATITUDE,
  LONGITUDE = LONGITUDE,
  RADIUS = RADIUS)

SBI.distance.like.crocs$STATIONNAME<-as.character(SBI.distance.like.crocs$STATIONNAME)

v.bass<-ReadInputData(infile=v.data.like.crocs,
                            iHoursToAdd=10,
                            dateformat = NULL,
                            sVemcoFormat='1.0')

GenerateAnimationKMLFile_Multitag(v.bass,
                                      SBI.distance,
                                      "Bass Moves.kml",
                                      sLocation= "STATIONNAME")


```

##Vtrack Animation of Track Lines

In the second file, you have to change the transmitter ID to be able to create different tracks for each fish
```{r}
TransmitterList <- unique(sbi.taginfo$tag_id) # Extract the transmitter names



GenerateAnimationKMLFile_Track(v.bass, # VTrack archive file
                               TransmitterList[2], # Transmitter id Change this number to generate new tag track
                               SBI.distance, # points file
                               "Track1.kml", # file name
                               "669900", # colour of the track
                               sLocation="STATIONNAME")
```

#Social Networks

This code is adapted from a work flow developed by Alyssa Clevenstine for her GSB research off Catalina. Because GSB are aggregate spawners, Betweenness and Closeness Centrality could be indicative of spawning aggregations.

```{r}

#create data.frame with necessary data
GSB.net<-v.data %>% transmute(
  Date_Time_PST = local,
  Date_Time_PST_Rounded = as.POSIXct(paste(substring(local, 1,13), "00:00", sep = ":")),
  Site = Station.Name,
  Transmitter = Transmitter,
  ID = substring(Transmitter, 10))

#subset summer dates
summer.19.GSB<-GSB.net[which(GSB.net$Date_Time_PST_Rounded > "2019-06-01 00:00:00"),]

#create week number
summer.19.GSB2<-summer.19.GSB %>% mutate(week = (year(Date_Time_PST_Rounded) - year(min(Date_Time_PST_Rounded)))*52 + week(Date_Time_PST_Rounded) - week(min(Date_Time_PST_Rounded)), week2 = (as.numeric(Date_Time_PST_Rounded) %/% 7) - (as.numeric(min(Date_Time_PST_Rounded)) %/% 7)) %>%
  arrange(Date_Time_PST_Rounded)

#change ID to a factor
tagIDs<-as.factor(sbi.taginfo$tag_id)

# create Julian Time Stamp since 1970
summer.19.GSB2$Date_Time_PST_Julian<-as.POSIXct(summer.19.GSB2$Date_Time_PST_Rounded, tz = "America/Los Angeles", orgin = "1970-01-01")

summer.19.GSB2$TimeStamp<-as.numeric(unclass(summer.19.GSB2$Date_Time_PST_Julian))

All.Summer.Network <- asnipe::gmmevents(time = summer.19.GSB2$TimeStamp, identity = summer.19.GSB2$ID, location = summer.19.GSB2$Site, global_ids = tagIDs, verbose = TRUE, splitGroups = TRUE)

summer.gbi <- All.Summer.Network$gbi
summer.events <- All.Summer.Network$metadata
summer.obs.per.event <- All.Summer.Network$B

sum.network <- get_network(summer.gbi, data_format = "GBI", association_index = "SRI") #calculate a network from
#a group by individual matrix and set the assoc. index to "simple ratio index"

net <- graph.adjacency(sum.network, mode = "undirected", diag = FALSE, weighted = TRUE)
#turn it into an igraph

dg.wtd <- graph.strength(net) #calculate the degree of the network

dg.wtd <- graph.strength(net) #calculate the degree of the network

btwn <- igraph::betweenness(net, v = V(net), directed = FALSE, weights = NULL, nobigint = TRUE, normalized = FALSE)
closeness.cent <- igraph::closeness(net, mode="all")
dg.cent <- centr_degree(net, mode = "all")


V(net)$size <- igraph::degree(net, v = V(net))
E(net)$width <- E(net)$weight

plot(net, vertex.size = igraph::degree(net, v = V(net))*.5,
     vertex.label.cex = 1.5, vertex.label.dist = 1, vertex.label.color = "black",
     vertex.color = "orange", edge.width = E(net)$weight, edge.arrow.size = 0.1, edge.curved = 0.1)
title(main = "Whole Summer", cex.main = 2)


summer.gbi <- as.data.frame(All.Summer.Network$gbi)
summer.events <- as.data.frame(All.Summer.Network$metadata)
summer.obs.per.event <- as.data.frame(All.Summer.Network$B)

```

Here is the while loop to run the network by week
```{r}
datalist = list()
for (i in levels(summer.19.GSB2$week < 21)) {
x<-count(summer.19.GSB2$ID)
x$i<-i
datalist[[i]]<-x
}

```


##Residency Metric Creation
```{r}
b.res <- RunResidenceExtraction(v.bass, "STATIONNAME", iResidenceThreshold = 3, iTimeThreshold = 60, sDistanceMatrix = DirectDM)

```




## Map raw detection patterns

```{r}
ATTdata$Tag.Detections %>% 
  group_by(Transmitter, Station.Name, Longitude, Latitude) %>% 
  summarise(num.det=n()) %>% 
  st_as_sf(coords=c("Longitude","Latitude"), crs=4326) %>% 
  mapview(zcol = "Transmitter", burst=T, legend=F, cex = "num.det") +
  ATTdata$Station.Information %>% 
  st_as_sf(coords=c("Station.Longitude", "Station.Latitude"), crs=4326) %>% 
  mapview(alpha.regions = 0, color = "white", homebutton=F, legend = F, layer.name = "Receivers")
```











